### 2020-01-16 ISSTA/conclusions.tex
shrink conc

Unexpected nondeterminism of software systems frustrates users,
whether they be humans or (more importantly) other software systems.
Nondeterminism is even more pernicious in software testing,
frustrating debugging efforts (Heisenbugs \cite{Heisenbug} and
Mandelbugs \cite{GrottkeBugs,FaultTriggers} are widely loathed), and
leading to the costly problem of flaky tests
\cite{miccoflaky,listfieldtestanalysis}.
This paper proposes a formulation of types of nondeterminism, and a
practical approach to using automated test generation to detect
nondeterminism, especially in the library code that underlies most
systems. In addition to traditional \emph{horizontal} nondeterminism
(the phenomenon behind Mandelbugs and flaky tests), we also discuss
the related concept of \emph{vertical} nondeterminism, which is more
frequently simply a software bug.  We implemented our approach in the
TSTL automated test generation system for Python, and demonstrated the
simplicity and utility of the approach on real-world examples.

As future work, we would like to make the automatic detection of
values that should not be visible (e.g., count towards nondeterminism)
possible: if a value (e.g., a timestamp) is always different in every
run, it is likely opaque.  This would make
testing libraries mixing deterministic behavior and nondeterministic
behavior, e.g. cryptographic librarie, much easier. Similarly, in order to extend the
utility of vertical nondeterminism, we would like to
automatically identify usually-idempotent operations.
%We would also like to lower the cost
%of checking process nondeterminism, perhaps by using methods taken in
%AFL \cite{aflfuzz} to avoid high startup costs for test executions.
Finally, we are interested in using test decomposition
\cite{Composition} to more easily isolate, understand, and avoid
nondeterminism in tests, and to mitigate flaky tests.
Reliable detection of nondeterminism is a first step towards
evaluating such fine-grained proactive flaky test mitigations.

>>>>

Unexpected nondeterminism of software systems frustrates users,
whether they be humans or (more importantly) other software systems.
Nondeterminism is even more pernicious in software testing,
frustrating debugging efforts (Heisenbugs \cite{Heisenbug} and
Mandelbugs \cite{GrottkeBugs,FaultTriggers} are widely loathed), and
leading to the costly problem of flaky tests
\cite{miccoflaky,listfieldtestanalysis}.
This paper proposes a formulation of types of nondeterminism, and a
practical approach to using automated test generation to detect
nondeterminism, especially in the library code that underlies most
systems. In addition to traditional \emph{horizontal} nondeterminism, we introduce
the concept of \emph{vertical} nondeterminism bugs.  
\begin{comment}
We implemented our approach in the
TSTL automated test generation system for Python, and demonstrated the
simplicity and utility of the approach on real-world examples.
\end{comment}

As future work, we would like to support the automatic detection of opaque
values; if a value (e.g., a timestamp) is different in \emph{every}
trace, it is likely opaque.  This would make
testing libraries mixing deterministic and nondeterministic
behavior, e.g. cryptographic libraries, much easier. Similarly, for
vertical nondeterminism, we would like to
automatically identify usually-idempotent operations.
%We would also like to lower the cost
%of checking process nondeterminism, perhaps by using methods taken in
%AFL \cite{aflfuzz} to avoid high startup costs for test executions.
Finally, we are interested in using test decomposition
\cite{Composition} to more isolate, understand, and mitigate
nondeterminism in unit tests, and to mitigate flaky tests.
Reliable detection of nondeterminism is a first step towards
evaluating such fine-grained proactive flaky test mitigation.
